---
title: "Data Import and Processing"
author: "Marie, Anna, Michel"
date: today
format: 
    html:
      toc: true
      self-contained: true
      highlight-style: github
      code-line-numbers: true
editor_options: 
    chunk_output_type: inline
---

# Methods

The differential gene expression and gene set analyses were performed in RStudio (Posit team (2023), version 2023.6.2.561, R version 4.3.2 (2023-10-31)).

## Differential Transcript Usage Analysis

After detecting differentially expressed genes, differentially expressed isoforms were identified using Differential Transcript Usage (DTU) analysis pipeline using DRIMSeq and DEXSeq libraries. DTU complements DGE as it enables the identification of alternative splicing and isoform switches even if the gene expression does not change between conditions [1, 2]. As for DGE, the aim was to identify differentially expressed isoforms in samples generated over course of the 5 days and  to compare the results of the differential gene expression analysis of the Illumina+Salmon data (short read data) with the results of the PacBio+Bambu data (long read data). To do so, in total 8 different count data frames were analyzed. For each of the 4 NDR cutoffs (0.025, 0.05, 0.1, 0.2), there are two different count data frames, one derived from Illumina+Salmon and one derived from PacBio+Bambu. **I didn't change this - I think we can unite DGE and DTU and write that both were conducted - the first two sentences I wrote can just be added after the first sentence of Michel.**


The inspiration for this protocol was taken from:

https://ycl6.gitbook.io/guide-to-rna-seq-analysis/differential-expression-analysis/differential-transcript-usage/dtu-using-dexseq

https://www.bioconductor.org/packages/release/workflows/vignettes/rnaseqDTU/inst/doc/rnaseqDTU.html

# References

[1] Marques-Coelho, D., Iohan, L.d.C.C., Melo de Farias, A.R. et al. Differential transcript usage unravels gene expression alterations in Alzheimer’s disease human brains. npj Aging Mech Dis 7, 2 (2021). https://doi.org/10.1038/s41514-020-00052-5

[2] https://www.bioconductor.org/packages/release/workflows/vignettes/rnaseqDTU/inst/doc/rnaseqDTU.html

[3] Anders S, Reyes A, Huber W. Detecting differential usage of exons from RNA-seq data. Genome Res. 2012 Oct;22(10):2008-17. doi: 10.1101/gr.133744.111.

[4] Nowicka M, Robinson MD. DRIMSeq: a Dirichlet-multinomial framework for multivariate count outcomes in genomics. F1000Res. 2016 Jun 13;5:1356. doi: 10.12688/f1000research.8900.2.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(42)
```

## Load packages

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggvenn)
library(DRIMSeq)
library(DEXSeq)
library(stageR)
library(stringr)
library(ggplot2)
library(ggbeeswarm)
library(UpSetR)
library(viridis)
```

## Load data

The data is loaded from RDS-objects that have been created with the import_processing.qmd file. For each of the 4 NDR cutoffs (0.025, 0.05, 0.1, 0.2), two data frames exist containing the expression data (counts) of the two different technologies used to obtain the data:

    Illumina sequencing and Salmon quantification (short read)
    PacBio sequencing and Bambu quantification (long read)

In total, this adds up to 8 different data frames. Each of these data frames contains 11 samples, which are replicates from one of five different time points. Each time point has 1-3 replicates.

### Load data
```{r}
df_list_bambu <- readRDS("./bambu.rds")
df_list_salmon <- readRDS("./salmon.rds")
df_list_meta <- readRDS("./meta.rds")

## Salmon
salmon_0.025 <- df_list_salmon$salmon_0.025
salmon_0.05 <- df_list_salmon$salmon_0.05
salmon_0.1 <- df_list_salmon$salmon_0.1
salmon_0.2 <- df_list_salmon$salmon_0.2
## Bambu
bambu_0.025 <- df_list_bambu$bambu_0.025
bambu_0.05 <- df_list_bambu$bambu_0.05
bambu_0.1 <- df_list_bambu$bambu_0.1
bambu_0.2 <- df_list_bambu$bambu_0.2
```

# Testing Pipeline

The workflow was taken from ![here](https://www.bioconductor.org/packages/release/workflows/vignettes/rnaseqDTU/inst/doc/rnaseqDTU.html) and tested on the given dataset. The workflow makes use of 3 libraries vignettes: DRIMSeq, DEXSeq and stageR

DEXSeq assumes that the feature (transcript) counts follow a negative binomial (NB) distribution and considers them as relative proportions of the group (the gene) using an interaction term in a generalized linear model (GLM) [2, 3]. DRIMSeq assumes that feature proportions follow the Dirichlet distribution (an Dirichlet Multinomial model (MD) for each gene) where the total count for the gene is considered fixed [2, 4]. 

## salmon_0.025

Here, I considered using `tximport` library for salmon data, however, I realized that it has not been adapted for Bambu data and decided to leave it at counts (the same as in DGE).

```{r eval=F}
library(tximport)
txi <- tximport(c("/home/annatoidze/Documents/ETHZ/STA426/project/data/0.1_salmon_day0-rep1/quant.sf"), type="salmon", txOut=TRUE,
                countsFromAbundance="scaledTPM")
cts <- txi$counts
cts <- cts[rowSums(cts) > 0,]
patterns_to_remove <- c("^DQ", "^EF", "^SIRV")
cts <- data.frame(cts)
cts <-
  cts[!grepl(paste(patterns_to_remove, collapse = "|"),
                      rownames(cts)), , drop =F]
```

It is checked how many reads (the range) that mapped to transcriptome using salmon. Also, `txdf` dataframe is created which maps transcripts to corresponding genes.

```{r}
range(colSums(salmon_0.025)/1e6) # Between 14.24 and 93.31 million reads mapped to transcriptome based on the replicate
# Transcript to gene mappings
txdf <- df_list_meta$metagenes_0.025
all(rownames(salmon_0.025) %in% rownames(txdf))
```

The `samps` dataframe relates the sample identifiers to the conditions (in this case, day of the experiment). The `cts` dataframe corresponds to the counts. In order to run DRIMSeq, a dataframe `counts` with the gene ID, the feature (transcript) ID, and then columns for each of the samples is built:

```{r}
samps <- df_list_meta$metadata
samps <- samps %>% column_to_rownames(var="sampleID")
colnames(samps) <- c("sample_id", "day")
samps$day <- factor(samps$day)
cts <- salmon_0.025
colnames(cts) <- samps$sample_id
counts <- data.frame(gene_id=txdf$gene_id,
                     feature_id=rownames(txdf),
                     cts)
```

### DRIMSeq

We create a `dmDSdata` object with previously generated dataframes.

```{r}
d <- dmDSdata(counts=counts, samples=samps)
d
```

As the rows of the object are gene oriented, the genes can by checked by rows (sometimes for than one transcript mapped, thus >1 rows shown even if selecting just one):

```{r}
counts(d[1,])
```

### Filtering

The filtering is done using the function `dmFilter` from DRIMSeq package. *n* is defined as the minimal number of samples where the genes should be expressed. *n.small* is the minimal number of samples where the features should be expressed. Both are set to 3, as that is the number of replicates for day 0 and 5. Additional filters are also added: for a transcript to be retained in the dataset, it is required that (1) it has a count of at least 10 in at least *n.small* samples, (2) it has a relative abundance proportion of at least 0.1 in at least *n.small* samples, and (3) the total count of the corresponding gene is at least 10 in all *n* samples.

```{r}
n <- 3
n.small <- 3
d <- dmFilter(d,
              min_samps_feature_expr=n.small, min_feature_expr=10,
              min_samps_feature_prop=n.small, min_feature_prop=0.1,
              min_samps_gene_expr=n, min_gene_expr=10)
d
plotData(d)
```

It can be seen how many of the remaining genes have *N* isoforms.

```{r}
table(table(counts(d)$gene_id))
```

### Creating design matrix

The design matrix is created in a way that the first time point, day0, is used as a baseline (intercept of the model).

```{r}
design_full <- model.matrix(~day, data=DRIMSeq::samples(d))
colnames(design_full)
```

### Precision calculation

`dmPrecision` is used t estimate the precision by maximum likelihood. It is inversely related to the dispersion in the DM model via following relation: $dispersion = \frac{1}{1 + precision}$. `dmFit` is used to fit regression coefficients and estimate feature propotrions for each sample. `dmTest` is used to perform null hypothesis likelihood ratio testing on `day5`. `plotPrecision` is used to plot precision (y-axis) against mean expression (x-axis). In RNA-seq data there is usually a trend of decrease of dispersion (here, increase of precision) with increase of mean expression. This is the trend observed here.

```{r}
system.time({
  d <- dmPrecision(d, design=design_full)
  d <- dmFit(d, design=design_full)
})

plotPrecision(d)
# Numeric vector of mean gene expression.
head(mean_expression(d))
# Numeric value of estimated common precision.
common_precision(d)
# Numeric vector of estimated gene-wise precisions.
head(genewise_precision(d))

## Get fitted proportions - estimated feature ratios for each sample
head(proportions(d))
## Get the DM regression coefficients (gene-level) 
head(coefficients(d))
## Get the BB regression coefficients (feature-level) 
head(coefficients(d), level = "feature")
```


### Testing and results

```{r}
## Fit null model proportions and perform the LR test to detect DTU
d <- dmTest(d, coef="day5")

## Plot the gene-level p-values
plotPValues(d)

## Get the gene-level results
## Plot feature proportions for a top DTU gene
res <- DRIMSeq::results(d)
res <- res[order(res$pvalue, decreasing = FALSE), ]

### Get the feature-level results
res.txp <- DRIMSeq::results(d, level="feature")
res.txp <- res.txp[order(res.txp$pvalue, decreasing = FALSE), ]

# Significant gene results
idx <- which(res$adj_pvalue < 0.05)
res[idx,]
length(idx)

# Significant isoform results
idx.txp <- which(res.txp$adj_pvalue < 0.05)
res.txp[idx.txp,]
length(idx.txp)
```

*pvalue* may contain NA values which may cause downstream errors (in stagewise analysis by `stageR`) so they are converted to 1`s.

```{r}
no.na <- function(x) ifelse(is.na(x), 1, x)
res$pvalue <- no.na(res$pvalue)
res.txp$pvalue <- no.na(res.txp$pvalue)
```


### Plotting

For a given gene, plot the observed and estimated with Dirichlet-multinomial model feature proportions in each group. Estimated group proportions are marked with diamond shapes.


```{r}
top_gene_id <- res$gene_id[1]
plotProportions(d, gene_id = top_gene_id, group_variable = "day")

plotProportions(d, gene_id = top_gene_id, group_variable = "day", 
  plot_type = "lineplot")

plotProportions(d, gene_id = top_gene_id, group_variable = "day", 
  plot_type = "ribbonplot")
```

### StageR following DRIMSeq

stageR package does stage-wise analysis of data, by screening first at the gene level for evidence of DTU (screening stage) and confirms which trancsripts within those significant genes show evidence of DTU (confirmation stage).

```{r}
nrow(res)
```

A vector of per-gene p-values `pScreen` for the screening stage is constructed. Addiionally, the gene and transcript version numbers are stripped from their Ensembl IDs (only first 15 characters are kept).

```{r}
pScreen <- res$pvalue
strp <- function(x) substr(x,1,15)
names(pScreen) <- strp(res$gene_id)
```

A one column matrix of the confirmation p-values `pConfirmation` is constructed:

```{r}
pConfirmation <- matrix(res.txp$pvalue, ncol=1)
rownames(pConfirmation) <- strp(res.txp$feature_id)
```

A two column dataframe `tx2gene`with the transcript and gene identifiers is arranged:

```{r}
tx2gene <- res.txp[,c("feature_id", "gene_id")]
for (i in 1:2) tx2gene[,i] <- strp(tx2gene[,i])
```

`stageR` analysis is perform with *overall false discovery rate* (OFDR, alpha parameter) equal to 0.05. 

```{r}
stageRObj <- stageRTx(pScreen=pScreen, pConfirmation=pConfirmation,
                      pScreenAdjusted=FALSE, tx2gene=tx2gene)
stageRObj <- stageWiseAdjustment(stageRObj, method="dtu", alpha=0.05)
suppressWarnings({
  drim.padj <- getAdjustedPValues(stageRObj, order=FALSE,
                                  onlySignificantGenes=TRUE)
})
head(drim.padj)
nrow(drim.padj)
```

The data.frame dex.padj summarizes the adjusted p-values from the two-stage analysis. Only genes that passed the filter (screening) are included in the table. The transcripts (column `trancsript`) with values less than 0.05 (5 \% OFDR) pass the confirmation stage.

## DEXSeq

The filtered DRIMSeq object is used for DEXSeq workflow, as the filtering is easier, and the analysis is sped up. `DEXSeqDataSet` is built from `d`, filtered DRIMseq object, sample information (`sample.data`), a design formula, transcript ID and gene ID:

```{r}
sample.data <- DRIMSeq::samples(d)
# We start with already filtered DRIMSeq object
count.data <- round(as.matrix(counts(d)[,-c(1:2)]))
dxd <- DEXSeqDataSet(countData=count.data,
                     sampleData=sample.data,
                     design=~sample + exon + day:exon,
                     featureID=counts(d)$feature_id,
                     groupID=counts(d)$gene_id)
```

DEXSeq analysis is performed using the following 3 functions: (1) estimation of size factors using `estimateSizeFactors`, (2) estimation of dispersion `estimateDispersions`, then (3) perform a likelihood ratio test for differential exon usage (in this case, transcripts) `testForDEU`. 

```{r}
system.time({
  dxd <- estimateSizeFactors(dxd)
  dxd <- estimateDispersions(dxd, quiet=TRUE)
  dxd <- testForDEU(dxd, reducedModel=~sample + exon)
})
```

The results tables (log2 fold changes and p-values) can be generated using the `DEXSeqResults `function and `perGeneQValue` function is used to compute a per-gene adjusted p-value.

```{r}
dxr <- DEXSeqResults(dxd, independentFiltering=FALSE)
qval <- perGeneQValue(dxr)
# Per gene
dxr.g <- data.frame(gene=names(qval),qval)
# Per transcript
dxr.t = as.data.frame(dxr[, c("featureID","groupID","pvalue")])
head(dxr.g)
```

Results:
```{r}
# Number of identified genes showing evidence for DTU
nrow(dxr.g[dxr.g$qval < 0.05,])
# Number of transcripts involved
nrow(dxr[dxr$padj < 0.05,])
```

### stageR following DEXSeq

The code for the procedure is similar to DRIMSeq procedure, with `alpha=0.05`.
```{r}
strp <- function(x) substr(x,1,15)
pConfirmation <- matrix(dxr$pvalue,ncol=1)
dimnames(pConfirmation) <- list(strp(dxr$featureID),"transcript")
pScreen <- qval
names(pScreen) <- strp(names(pScreen))
tx2gene <-  data.frame(dxr.t[,c("featureID", "groupID")], 
dxr.t[,c("featureID", "groupID")])
for (i in 1:2) tx2gene[,i] = strp(tx2gene[,i])
stageRObj <- stageRTx(pScreen=pScreen, pConfirmation=pConfirmation,
                      pScreenAdjusted=TRUE, tx2gene=tx2gene[1:2])
stageRObj = stageRTx(pScreen = pScreen, 
pConfirmation = pConfirmation, 
pScreenAdjusted = TRUE, 
tx2gene = tx2gene[1:2])

stageRObj = stageWiseAdjustment(stageRObj, method = "dtu", alpha = 0.05)

suppressWarnings({
  dex.padj <- getAdjustedPValues(stageRObj, order=FALSE,
                                 onlySignificantGenes=TRUE)
})

dex.padj = merge(tx2gene, dex.padj, by.x = c("groupID","featureID"), by.y = c("geneID","txID"))

head(dex.padj)
```

```{r}
length(unique(dex.padj[dex.padj$gene < 0.05,]$groupID))
table(dex.padj$transcript < 0.05)
```

### Exporting the results
```{r}
dex.norm = cbind(as.data.frame(stringr::str_split_fixed(rownames(counts(dxd)), ":", 2)), as.data.frame(counts(dxd, normalized = TRUE))[,1:20])
colnames(dex.norm) = c("groupID", "featureID", as.character(colData(dxd)$sample_id)[1:20])
row.names(dex.norm) = NULL
# Per-group normalised mean
dex.mean = as.data.frame(sapply( levels(samps$day), 
function(lvl) rowMeans(dex.norm[, 3:ncol(dex.norm)][, samps$day == lvl, drop = FALSE]) ))
colnames(dex.mean) <- paste0("mean.day", colnames(dex.mean))

# log2 fold change in expression
dex.log2fc = log2(dex.mean[2]/dex.mean[1])
colnames(dex.log2fc) = "log2fc"
rownames(dex.log2fc) = dex.norm$featureID

# Merge to create result data
dexData = cbind(dex.norm[,1:2], dex.mean, dex.norm[, 3:ncol(dex.norm)])
colnames(dexData)[1:2] <-  c("GeneID","TranscriptID")
dexData = dexData[order(dexData$GeneID, dexData$TranscriptID),]

# Merge to create result data
dexDTU = merge(dex.padj[,c("featureID.1","groupID.1","gene","transcript")], dex.log2fc, by.x = "featureID.1", by.y = "row.names")
colnames(dexDTU)[1:2] <-  c("TranscriptID","GeneID")
dexDTU = dexDTU[order(dexDTU$GeneID, dexDTU$TranscriptID),]

# Write results to files
write.table(dexData, file="DTU_DEXSeq-stageR_means_and_counts.txt", sep = "\t", quote = F, row.names = F, col.names = T)
write.table(dexDTU, file="DTU_DEXSeq-stageR_results.txt", sep = "\t", quote = F, row.names = F, col.names = T)
```


### Plotting

A custom function `plotExpression` was made to plot the transcript expression of a given gene as from ![here](https://ycl6.gitbook.io/guide-to-rna-seq-analysis/differential-expression-analysis/differential-transcript-usage/dtu-using-dexseq).

```{r}
# We will use this function in both DRIMSeq & DEXSeq workflows
plotExpression <- function(expData = NULL, geneID = NULL, samps = NULL, isProportion = FALSE) {
        colnames(expData)[1:2] = c("gid","tid")
        sub = subset(expData, gid == geneID)
        print(sub)
        sub = reshape2::melt(sub, id = c("gid", "tid"))
        sub = merge(samps, sub, by.x = "sample_id", by.y = "variable")
        print(sub)
        if(!isProportion) {
                sub$value = log(sub$value)
        }
        print(sub)

        clrs = c("dodgerblue3", "maroon2",  "forestgreen", "darkorange1", "blueviolet", "firebrick2",
"deepskyblue", "orchid2", "chartreuse3", "gold", "slateblue1", "tomato" , "blue", "magenta", "green3",
"yellow", "purple3", "red" ,"darkslategray1", "lightpink1", "lightgreen", "khaki1", "plum3", "salmon")

        p = ggplot(sub, aes(tid, value, color = day, fill = day)) +
        geom_boxplot(alpha = 0.4, outlier.shape = NA, width = 0.8, lwd = 0.5) +
        stat_summary(fun = mean, geom = "point", color = "black", shape = 5, size = 3, position=position_dodge(width = 0.8)) +
        scale_color_manual(values = clrs) + scale_fill_manual(values = clrs) +
        geom_quasirandom(color = "black", size = 1, dodge.width = 0.8) + theme_bw() +
        ggtitle(geneID) + xlab("Transcripts")

        if(!isProportion) {
                p = p + ylab("log(Expression)")
        } else {
                p = p + ylab("Proportions")
        }
        p
}

```


```{r}
# Plot the normalised counts for one of the significant genes, where we can see evidence of switching
gene_id <-  unique(dex.padj[order(dex.padj$transcript, dex.padj$gene),]$groupID.1)[1]
plotExpression(dex.norm, gene_id, samps, isProportion = T)
```
## Comparison of DRIMSeq and DEXSeq

Number of identified transcripts from DEXSeq: 

```{r}
dim(dex.padj[order(dex.padj$gene, decreasing = T), ])
```

Number of identified transcripts from DRIMSeq:

```{r}
dim(drim.padj[order(drim.padj$gene, decreasing = T), ])
```


Check the intersection between the different methods for screened genes:

```{r}
length(intersect(dex.padj$groupID, drim.padj$geneID))
venn <- list(DEX = dex.padj$groupID, DRIM = drim.padj$geneID)
ggvenn(venn)
```

Check the intersection between the different methods for features:

```{r}
length(intersect(dex.padj$featureID, drim.padj$txID))
venn <- list(DEX = dex.padj$featureID, DRIM = drim.padj$txID)
ggvenn(venn)
```

# Making a function for the pipeline using DEXSeq x DRIMSeq

```{r}
samps <- df_list_meta$metadata
samps <- samps %>% column_to_rownames(var="sampleID")
colnames(samps) <- c("sample_id", "day")
samps$day <- factor(samps$day)

DTU <- function(salmon, txdf, sample_name, n = 11, n.small = 11){
  # DRIMSeq Filtering
  cts <- salmon
  colnames(cts) <- samps$sample_id
  counts <- data.frame(gene_id=txdf$gene_id,
                     feature_id=rownames(txdf),
                     cts)
  d <- dmDSdata(counts=counts, samples=samps)
  d <- dmFilter(d,
                min_samps_feature_expr=n.small, min_feature_expr=10,
                min_samps_feature_prop=n.small, min_feature_prop=0.1,
                min_samps_gene_expr=n, min_gene_expr=10)
  print(plotData(d))
  png(paste0(sample_name, "plotData.DEXSeq-stageR.png"), width=6, height=6, units = "in", res = 300)
  print(plotData(d))
  dev.off()
  print(table(table(counts(d)$gene_id)))
  sample.data <- DRIMSeq::samples(d)
  
  # DEXSeq
  count.data <- round(as.matrix(counts(d)[,-c(1:2)]))
  dxd <- DEXSeqDataSet(countData=count.data,
                       sampleData=sample.data,
                       design=~sample + exon + day:exon,
                       featureID=counts(d)$feature_id,
                       groupID=counts(d)$gene_id)
  dxd <- estimateSizeFactors(dxd)
  dxd <- estimateDispersions(dxd, quiet=TRUE)
  dxd <- testForDEU(dxd, reducedModel=~sample + exon)
  dxr <- DEXSeqResults(dxd, independentFiltering=FALSE)
  qval <- perGeneQValue(dxr)
  # Per gene
  dxr.g <- data.frame(gene=names(qval),qval)
  # Per transcript
  dxr.t = as.data.frame(dxr[, c("featureID","groupID","pvalue")])
  # Number of identified genes showing evidence for DTU
  print(nrow(dxr.g[dxr.g$qval < 0.05,]))
  # Number of transcripts involved
  print(nrow(dxr[dxr$padj < 0.05,]))

  # stageR
  strp <- function(x) substr(x,1,15)
  pConfirmation <- matrix(dxr$pvalue,ncol=1)
  dimnames(pConfirmation) <- list(strp(dxr$featureID),"transcript")
  pScreen <- qval
  names(pScreen) <- strp(names(pScreen))
  tx2gene <-  data.frame(dxr.t[,c("featureID", "groupID")], 
  dxr.t[,c("featureID", "groupID")])
  for (i in 1:2) tx2gene[,i] = strp(tx2gene[,i])
  stageRObj <- stageRTx(pScreen=pScreen, pConfirmation=pConfirmation,
                        pScreenAdjusted=TRUE, tx2gene=tx2gene[1:2])
  stageRObj = stageRTx(pScreen = pScreen, 
  pConfirmation = pConfirmation, 
  pScreenAdjusted = TRUE, 
  tx2gene = tx2gene[1:2])
  
  stageRObj = stageWiseAdjustment(stageRObj, method = "dtu", alpha = 0.05)
  
  suppressWarnings({
    dex.padj <- getAdjustedPValues(stageRObj, order=FALSE,
                                   onlySignificantGenes=TRUE)
  })
  
  dex.padj = merge(tx2gene, dex.padj, by.x = c("groupID","featureID"), by.y = c("geneID","txID"))
  print(head(dex.padj[order(dex.padj$gene, decreasing = T), ]))
  
  # Normalize counts
  dex.norm = cbind(as.data.frame(stringr::str_split_fixed(rownames(counts(dxd)), ":", 2)), as.data.frame(counts(dxd, normalized = TRUE))[,1:20])
  colnames(dex.norm) = c("groupID", "featureID", as.character(colData(dxd)$sample_id)[1:20])
  row.names(dex.norm) = NULL

  # Per-group normalised mean
  dex.mean = as.data.frame(sapply( levels(samps$day), 
  function(lvl) rowMeans(dex.norm[, 3:ncol(dex.norm)][, samps$day == lvl, drop = FALSE]) ))
  colnames(dex.mean) <- paste0("mean.day", colnames(dex.mean))
  
  # log2 fold change in expression
  dex.log2fc = log2(dex.mean[2]/dex.mean[1])
  colnames(dex.log2fc) = "log2fc"
  rownames(dex.log2fc) = dex.norm$featureID
  
  # Merge to create result data
  dexData = cbind(dex.norm[,1:2], dex.mean, dex.norm[, 3:ncol(dex.norm)])
  colnames(dexData)[1:2] <-  c("GeneID","TranscriptID")
  dexData = dexData[order(dexData$GeneID, dexData$TranscriptID),]
  
  # Merge to create result data
  dexDTU = merge(dex.padj[,c("featureID.1","groupID.1","gene","transcript")], dex.log2fc, by.x = "featureID.1", by.y = "row.names")
  colnames(dexDTU)[1:2] <-  c("TranscriptID","GeneID")
  dexDTU = dexDTU[order(dexDTU$GeneID, dexDTU$TranscriptID),]
  
  # Write results to files
  write.table(dexData, file=paste0(sample_name, "DTU_DEXSeq-stageR_means_and_counts.txt"), sep = "\t", quote = F, row.names = F, col.names = T)
  write.table(dexDTU, file=paste0(sample_name, "DTU_DEXSeq-stageR_results.txt"), sep = "\t", quote = F, row.names = F, col.names = T)

  # Plot the normalised counts for one of the significant genes, where we can see evidence of switching
  gene_id <-  unique(dex.padj[order(dex.padj$transcript, dex.padj$gene),]$groupID.1)[1]
  
  print(plotExpression(dex.norm, gene_id, samps, isProportion = T))
  png(paste0(sample_name, "plotExpression-stageR.png"), width=6, height=6, units = "in", res = 300)
  print(plotExpression(dex.norm, gene_id, samps, isProportion = T))
  dev.off()
  
  return(dexDTU)
}
```

### NDR 0.025
```{r}
# Transcript to gene mappings
txdf <- df_list_meta$metagenes_0.025
res_salmon_0.025 <- DTU(salmon_0.025, txdf, "salmon_0.025.", n = 3, n.small = 3)
res_bambu_0.025 <- DTU(bambu_0.025, txdf,  "bambu_0.025.", n = 3, n.small = 3)

confirmed_salmon_0.025 <- res_salmon_0.025[res_salmon_0.025$transcript < 0.05, ]
nrow(confirmed_salmon_0.025)
confirmed_bambu_0.025 <- res_bambu_0.025[res_bambu_0.025$transcript < 0.05, ]
nrow(confirmed_bambu_0.025)
```

For Illumina+Salmon, out of the 9901 tested genes and 27084 tested features, DEXSeq test identified 1457 genes showing evidence of isoform switching involving 2409 transcripts. For PacBio+Bambu out of the 7088 tested genes and 17734 tested features, 3166 genes involving 6967 trancsripts were identified.

After the stageR procedure, 2358 transcripts passed the confirmation stage for Illumina+Salmon and 6316 transcripts for PacBio+Bambu on a target 5% overall false discovery rate (OFDR). This means that, in expectation, no more than 5% of the genes that pass screening will either (1) not contain any DTU, so be falsely screened genes, or (2) contain a falsely confirmed transcript. 

### Upset Plot
```{r}
# Screened Genes
length(intersect(confirmed_salmon_0.025$GeneID, confirmed_bambu_0.025$GeneID))
venn <- list(SALMON = confirmed_salmon_0.025$GeneID, BAMBU = confirmed_bambu_0.025$GeneID)
ggvenn(venn)
upset(fromList(venn))

# Confirmed Transcripts
length(intersect(confirmed_salmon_0.025$TranscriptID, confirmed_bambu_0.025$TranscriptID))
venn <- list(SALMON = confirmed_salmon_0.025$TranscriptID, BAMBU = confirmed_bambu_0.025$TranscriptID)
ggvenn(venn)
upset(fromList(venn))
```

From the UpSet plot and Venn Diagram above it is visible that through PacBio+Bambu many more transcripts were detected through DTU. The overlap between Illumina+Salmon was also only roughly 11 \%.

### Comparsion of pAdj and logFC 
```{r}
both_detected <-
  intersect(res_bambu_0.025$TranscriptID, res_salmon_0.025$TranscriptID)
sorted_res_salmon_0.025 <- res_salmon_0.025[order(res_salmon_0.025$TranscriptID),]
both_sorted_res_salmon_0.025 <- sorted_res_salmon_0.025[which(sorted_res_salmon_0.025$TranscriptID %in% both_detected), ]
sorted_res_bambu_0.025 <- res_bambu_0.025[order(res_bambu_0.025$TranscriptID),]
both_sorted_res_bambu_0.025 <- sorted_res_bambu_0.025[which(sorted_res_bambu_0.025$TranscriptID %in% both_detected), ]

padj <- data.frame(Salmon = both_sorted_res_salmon_0.025$transcript, Bambu = both_sorted_res_bambu_0.025$transcript)
ggplot(data = padj, mapping = aes(x = -log10(Salmon), y = -log10(Bambu))) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of P-values (NDR=0.025)", x = "-log10 P-value (Salmon)", y = "-log10 P-value (Bambu)")

logfc <- data.frame(Salmon = both_sorted_res_salmon_0.025$log2fc, Bambu = both_sorted_res_bambu_0.025$log2fc)
ggplot(data = logfc, mapping = aes(x = Salmon, y = Bambu)) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of LogFC (NDR=0.025)", x = "LogFC (Salmon)", y = "LogFC (Bambu)")
```

The comparison of the adjusted p-values for transcripts shows low correlation between technologies especially for higher values. The logFC values seem to be, however, better correlated with some deviations for the extreme values.


### NDR 0.05
```{r}
# Transcript to gene mappings
txdf <- df_list_meta$metagenes_0.05
res_salmon_0.05 <- DTU(salmon_0.05, txdf, "salmon_0.05.", n = 3, n.small = 3)
res_bambu_0.05 <- DTU(bambu_0.05, txdf,  "bambu_0.05.", n = 3, n.small = 3)

confirmed_salmon_0.05 <- res_salmon_0.05[res_salmon_0.05$transcript < 0.05, ]
nrow(confirmed_salmon_0.05)
confirmed_bambu_0.05 <- res_bambu_0.05[res_bambu_0.05$transcript < 0.05, ]
nrow(confirmed_bambu_0.05)
```

For Illumina+Salmon, out of the 9958 tested genes and 27223 tested features, DEXSeq test identified 1468 genes showing evidence of isoform switching involving 2436 transcripts. For PacBio+Bambu out of the 7171 tested genes and 17949 tested features, 3217 genes involving 7087 trancsripts were identified.

After the stageR procedure, 2375 transcripts passed the confirmation stage for Illumina+Salmon and 6427 transcripts for PacBio+Bambu on a target 5% overall false discovery rate (OFDR). This means that, in expectation, no more than 5% of the genes that pass screening will either (1) not contain any DTU, so be falsely screened genes, or (2) contain a falsely confirmed transcript. 

### Upset Plot
```{r}
# Screened Genes
length(intersect(confirmed_salmon_0.05$GeneID, confirmed_bambu_0.05$GeneID))
venn <- list(SALMON = confirmed_salmon_0.05$GeneID, BAMBU = confirmed_bambu_0.05$GeneID)
ggvenn(venn)
upset(fromList(venn))

# Confirmed Transcripts
length(intersect(confirmed_salmon_0.05$TranscriptID, confirmed_bambu_0.05$TranscriptID))
venn <- list(SALMON = confirmed_salmon_0.05$TranscriptID, BAMBU = confirmed_bambu_0.05$TranscriptID)
ggvenn(venn)
upset(fromList(venn))
```
From the UpSet plot and Venn Diagram above it is visible that through PacBio+Bambu many more transcripts were detected through DTU. The overlap between Illumina+Salmon was also only roughly 11 \%.

### Comparsion of pAdj and logFC 
```{r}
both_detected <-
  intersect(res_bambu_0.05$TranscriptID, res_salmon_0.05$TranscriptID)
sorted_res_salmon_0.05 <- res_salmon_0.05[order(res_salmon_0.05$TranscriptID),]
both_sorted_res_salmon_0.05 <- sorted_res_salmon_0.05[which(sorted_res_salmon_0.05$TranscriptID %in% both_detected), ]
sorted_res_bambu_0.05 <- res_bambu_0.05[order(res_bambu_0.05$TranscriptID),]
both_sorted_res_bambu_0.05 <- sorted_res_bambu_0.05[which(sorted_res_bambu_0.05$TranscriptID %in% both_detected), ]

padj <- data.frame(Salmon = both_sorted_res_salmon_0.05$transcript, Bambu = both_sorted_res_bambu_0.05$transcript)
ggplot(data = padj, mapping = aes(x = -log10(Salmon), y = -log10(Bambu))) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of P-values (NDR=0.05)", x = "-log10 P-value (Salmon)", y = "-log10 P-value (Bambu)")

logfc <- data.frame(Salmon = both_sorted_res_salmon_0.05$log2fc, Bambu = both_sorted_res_bambu_0.05$log2fc)
ggplot(data = logfc, mapping = aes(x = Salmon, y = Bambu)) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of LogFC (NDR=0.05)", x = "LogFC (Salmon)", y = "LogFC (Bambu)")
```

The comparison of the adjusted p-values for transcripts shows low correlation between technologies especially for higher values. The logFC values seem to be, however, better correlated with some deviations for the extreme values.

### NDR 0.1
```{r}
# Transcript to gene mappings
txdf <- df_list_meta$metagenes_0.1
res_salmon_0.1 <- DTU(salmon_0.1, txdf, "salmon_0.1.", n = 3, n.small = 3)
res_bambu_0.1 <- DTU(bambu_0.1, txdf,  "bambu_0.1.", n = 3, n.small = 3)

confirmed_salmon_0.1 <- res_salmon_0.1[res_salmon_0.1$transcript < 0.05, ]
nrow(confirmed_salmon_0.1)
confirmed_bambu_0.1 <- res_bambu_0.1[res_bambu_0.1$transcript < 0.05, ]
nrow(confirmed_bambu_0.1)
```

For Illumina+Salmon, out of the 10052 tested genes and 27506 tested features, DEXSeq test identified 1507 genes showing evidence of isoform switching involving 2505 transcripts. For PacBio+Bambu out of the 7366 tested genes and 18437 tested features, 3316 genes involving 7332 trancsripts were identified.

After the stageR procedure, 2430 transcripts passed the confirmation stage for Illumina+Salmon and 6661 transcripts for PacBio+Bambu on a target 5% overall false discovery rate (OFDR). This means that, in expectation, no more than 5% of the genes that pass screening will either (1) not contain any DTU, so be falsely screened genes, or (2) contain a falsely confirmed transcript. 

### Upset Plot
```{r}
# Screened Genes
length(intersect(confirmed_salmon_0.1$GeneID, confirmed_bambu_0.1$GeneID))
venn <- list(SALMON = confirmed_salmon_0.1$GeneID, BAMBU = confirmed_bambu_0.1$GeneID)
ggvenn(venn)
upset(fromList(venn))

# Confirmed Transcripts
length(intersect(confirmed_salmon_0.1$TranscriptID, confirmed_bambu_0.1$TranscriptID))
venn <- list(SALMON = confirmed_salmon_0.1$TranscriptID, BAMBU = confirmed_bambu_0.1$TranscriptID)
ggvenn(venn)
upset(fromList(venn))
```

From the UpSet plot and Venn Diagram above it is visible that through PacBio+Bambu many more transcripts were detected through DTU. The overlap between Illumina+Salmon was also only roughly 11 \%.

### Comparsion of pAdj and logFC 
```{r}
both_detected <-
  intersect(res_bambu_0.1$TranscriptID, res_salmon_0.1$TranscriptID)
sorted_res_salmon_0.1 <- res_salmon_0.1[order(res_salmon_0.1$TranscriptID),]
both_sorted_res_salmon_0.1 <- sorted_res_salmon_0.1[which(sorted_res_salmon_0.1$TranscriptID %in% both_detected), ]
sorted_res_bambu_0.1 <- res_bambu_0.1[order(res_bambu_0.1$TranscriptID),]
both_sorted_res_bambu_0.1 <- sorted_res_bambu_0.1[which(sorted_res_bambu_0.1$TranscriptID %in% both_detected), ]

padj <- data.frame(Salmon = both_sorted_res_salmon_0.1$transcript, Bambu = both_sorted_res_bambu_0.1$transcript)
ggplot(data = padj, mapping = aes(x = -log10(Salmon), y = -log10(Bambu))) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of P-values (NDR=0.1)", x = "-log10 P-value (Salmon)", y = "-log10 P-value (Bambu)")

logfc <- data.frame(Salmon = both_sorted_res_salmon_0.1$log2fc, Bambu = both_sorted_res_bambu_0.1$log2fc)
ggplot(data = logfc, mapping = aes(x = Salmon, y = Bambu)) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of LogFC (NDR=0.1)", x = "LogFC (Salmon)", y = "LogFC (Bambu)")
```

The comparison of the adjusted p-values for transcripts shows low correlation between technologies especially for higher values. The logFC values seem to be, however, better correlated with some deviations for the extreme values.


### NDR 0.2
```{r}
# Transcript to gene mappings
txdf <- df_list_meta$metagenes_0.2
res_salmon_0.2 <- DTU(salmon_0.2, txdf, "salmon_0.2.", n = 3, n.small = 3)
res_bambu_0.2 <- DTU(bambu_0.2, txdf,  "bambu_0.2.", n = 3, n.small = 3)

confirmed_salmon_0.2 <- res_salmon_0.1[res_salmon_0.2$transcript < 0.05, ]
nrow(confirmed_salmon_0.2)
confirmed_bambu_0.2 <- res_bambu_0.2[res_bambu_0.2$transcript < 0.05, ]
nrow(confirmed_bambu_0.2)
```
For Illumina+Salmon, out of the 10364 tested genes and 28447 tested features, DEXSeq test identified 1578 genes showing evidence of isoform switching involving 2569 transcripts. For PacBio+Bambu out of the 7847 tested genes and 19816 tested features, 3530 genes involving 7799 trancsripts were identified.

After the stageR procedure, 2528 transcripts passed the confirmation stage for Illumina+Salmon and 7108 transcripts for PacBio+Bambu on a target 5% overall false discovery rate (OFDR). This means that, in expectation, no more than 5% of the genes that pass screening will either (1) not contain any DTU, so be falsely screened genes, or (2) contain a falsely confirmed transcript. 

### Upset Plot
```{r}
# Screened Genes
length(intersect(confirmed_salmon_0.2$GeneID, confirmed_bambu_0.2$GeneID))
venn <- list(SALMON = confirmed_salmon_0.2$GeneID, BAMBU = confirmed_bambu_0.2$GeneID)
ggvenn(venn)
upset(fromList(venn))

# Confirmed Transcripts
length(intersect(confirmed_salmon_0.2$TranscriptID, confirmed_bambu_0.2$TranscriptID))
venn <- list(SALMON = confirmed_salmon_0.2$TranscriptID, BAMBU = confirmed_bambu_0.2$TranscriptID)
ggvenn(venn)
upset(fromList(venn))
```

From the UpSet plot and Venn Diagram above it is visible that through PacBio+Bambu many more transcripts were detected through DTU. The overlap between Illumina+Salmon was also only roughly 7 \%. This percentage is much less than for the cutoffs from before.

### Comparsion of pAdj and logFC 
```{r}
both_detected <-
  intersect(res_bambu_0.2$TranscriptID, res_salmon_0.2$TranscriptID)
sorted_res_salmon_0.2 <- res_salmon_0.2[order(res_salmon_0.2$TranscriptID),]
both_sorted_res_salmon_0.2 <- sorted_res_salmon_0.2[which(sorted_res_salmon_0.2$TranscriptID %in% both_detected), ]
sorted_res_bambu_0.2 <- res_bambu_0.2[order(res_bambu_0.2$TranscriptID),]
both_sorted_res_bambu_0.2 <- sorted_res_bambu_0.2[which(sorted_res_bambu_0.2$TranscriptID %in% both_detected), ]

padj <- data.frame(Salmon = both_sorted_res_salmon_0.2$transcript, Bambu = both_sorted_res_bambu_0.2$transcript)
ggplot(data = padj, mapping = aes(x = -log10(Salmon), y = -log10(Bambu))) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of P-values (NDR=0.2)", x = "-log10 P-value (Salmon)", y = "-log10 P-value (Bambu)")

logfc <- data.frame(Salmon = both_sorted_res_salmon_0.2$log2fc, Bambu = both_sorted_res_bambu_0.2$log2fc)
ggplot(data = logfc, mapping = aes(x = Salmon, y = Bambu)) +
  geom_point() + 
  geom_smooth() +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis() +
  labs(title = "Comparison of LogFC (NDR=0.2)", x = "LogFC (Salmon)", y = "LogFC (Bambu)")
```

The comparison of the adjusted p-values for transcripts shows low correlation between technologies especially for higher values. The logFC values seem to be, however, better correlated with some deviations for the extreme values.
